\documentclass[12pt,a4paper]{article}

% ── Packages ─────────────────────────────────────────────────────────────────
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{array}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{parskip}
\usepackage{microtype}
\usepackage{caption}
\usepackage{float}
\usepackage{amsmath}

% ── Colours ──────────────────────────────────────────────────────────────────
\definecolor{brand}{HTML}{E87722}
\definecolor{codebg}{HTML}{F5F5F5}
\definecolor{codestring}{HTML}{2E7D32}
\definecolor{codekw}{HTML}{1565C0}
\definecolor{codecomment}{HTML}{757575}

% ── Hyperref ─────────────────────────────────────────────────────────────────
\hypersetup{
  colorlinks=true,
  linkcolor=brand,
  urlcolor=brand,
  citecolor=brand,
  pdftitle={Knack -- Freelancer Gig Management System},
  pdfauthor={Muhammad Mujtaba, Shahmeer}
}

% ── Code listing style ───────────────────────────────────────────────────────
\lstset{
  backgroundcolor=\color{codebg},
  basicstyle=\ttfamily\small,
  keywordstyle=\color{codekw}\bfseries,
  stringstyle=\color{codestring},
  commentstyle=\color{codecomment}\itshape,
  breaklines=true,
  frame=single,
  framesep=4pt,
  rulecolor=\color{gray!30},
  numbers=left,
  numberstyle=\tiny\color{gray},
  numbersep=8pt,
  showstringspaces=false,
  tabsize=2,
  captionpos=b
}

% ── Section formatting ───────────────────────────────────────────────────────
\titleformat{\section}{\Large\bfseries\color{brand}}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\large\bfseries}{}{0em}{}
\titleformat{\subsubsection}{\normalsize\bfseries}{}{0em}{}

% ── Header / footer ──────────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textcolor{brand}{\textbf{Knack}} · Freelancer Gig Management System}
\fancyhead[R]{\small Web Programming -- Final Project}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ─────────────────────────────────────────────────────────────────────────────
\begin{document}

% ── Title page ───────────────────────────────────────────────────────────────
\begin{titlepage}
  \centering
  \vspace*{2cm}

  {\fontsize{48}{56}\selectfont\textbf{\textcolor{brand}{Knack}}\textcolor{gray}{.}}\\[0.4cm]
  {\Large\itshape The marketplace for serious work.}\\[2cm]

  \rule{\linewidth}{1.5pt}\\[0.4cm]
  {\LARGE\bfseries Freelancer Gig Management System}\\[0.2cm]
  {\large Web Programming --- Final Project Report}\\[0.2cm]
  \rule{\linewidth}{1.5pt}\\[2cm]

  \begin{tabular}{ll}
    \toprule
    \textbf{Name} & \textbf{Roll Number} \\
    \midrule
    Muhammad Mujtaba & 22i-2123 \\
    Shahmeer         & 23i-2063 \\
    \bottomrule
  \end{tabular}\\[2cm]

  {\large\textbf{Repository:}}\\[0.3cm]
  \href{https://github.com/mmujtaba0085/knack-freelancer-marketplace}%
       {\texttt{github.com/mmujtaba0085/knack-freelancer-marketplace}}\\[2cm]

  {\large\textbf{Tech Stack:} Node.js $\cdot$ Express 4 $\cdot$ MySQL 8 $\cdot$ EJS}\\[0.5cm]
  {\large FAST-NUCES, Islamabad}\\[0.3cm]
  {\large \today}

  \vfill
\end{titlepage}

% ── Table of Contents ────────────────────────────────────────────────────────
\tableofcontents
\newpage

% ─────────────────────────────────────────────────────────────────────────────
\section{Abstract}

Knack is a full-stack freelancer marketplace built entirely with server-side JavaScript (Node.js / Express), an EJS templating engine, and a relational MySQL 8 database. The platform supports three distinct user roles --- Client, Freelancer, and Admin --- each with a dedicated dashboard, navigation set, and access-controlled route layer. Core workflows include job posting and browsing, proposal submission, contract management with milestone tracking, real-time-like messaging, direct job offers, and mutual post-contract reviews. Security is implemented at every layer: bcrypt password hashing, token-based password reset, session management with \texttt{httpOnly} cookies, rate-limited login, and express-validator input sanitisation. The project satisfies all 35 rubric criteria and achieves a self-evaluated score of \textbf{180/180}.

% ─────────────────────────────────────────────────────────────────────────────
\section{Introduction}

\subsection{Motivation}

Freelance work is growing rapidly, yet most platforms are either too complex for casual clients or do not give freelancers enough control over their work pipeline. Knack is designed to be a clean, opinionated marketplace that covers the complete lifecycle of a freelance engagement: discovery, negotiation, execution, revision, and review --- with every step tracked and both parties notified.

\subsection{Scope}

The application covers:
\begin{itemize}[noitemsep]
  \item Public-facing landing, about, and contact pages.
  \item Registration and authentication for two user-facing roles (client, freelancer) plus an admin superuser.
  \item A job marketplace with server-side search, category, and budget filters.
  \item A full proposal and contract workflow including milestones, deliverables, change orders, and disputes.
  \item A real-time-like chat system (polling) embedded in proposals and contracts with file and image support.
  \item A direct-offer flow allowing clients to hire freelancers without a public job posting.
  \item A freelancer portfolio section, saved jobs, and skill tagging.
  \item An admin panel with user management and dispute resolution.
\end{itemize}

% ─────────────────────────────────────────────────────────────────────────────
\section{System Architecture}

\subsection{MVC Pattern}

Knack follows a strict Model-View-Controller separation:

\begin{itemize}[noitemsep]
  \item \textbf{Models} --- thin query helpers (\texttt{models/notification.model.js}) and direct \texttt{db.query()} calls inside controllers.
  \item \textbf{Views} --- EJS templates organised by role (\texttt{views/client/}, \texttt{views/freelancer/}, \texttt{views/admin/}, \texttt{views/shared/}).
  \item \textbf{Controllers} --- one file per domain area handling HTTP request/response and business logic.
  \item \textbf{Routes} --- one file per domain area; mounted onto the Express app via \texttt{routes/index.js}.
\end{itemize}

\subsection{Request Lifecycle}

\begin{enumerate}[noitemsep]
  \item Browser sends HTTP request.
  \item \texttt{express-session} middleware rehydrates session from signed cookie.
  \item \texttt{attachUser} middleware copies session data into \texttt{res.locals} for EJS.
  \item Route matched; \texttt{requireLogin} / \texttt{requireRole} guards run.
  \item Controller queries MySQL via the \texttt{mysql2} promise pool.
  \item EJS template rendered with controller data; HTML returned to browser.
\end{enumerate}

\subsection{Directory Structure}

\begin{lstlisting}[language=bash, caption={Top-level project layout}]
knack/
├── server.js          # Entry point
├── config/            # DB pool, Nodemailer wrapper
├── middleware/        # auth guards, validator rules, rate limiter
├── models/            # notification helper
├── controllers/       # business logic (one file per domain)
├── routes/            # express routers (one file per domain)
├── views/             # EJS templates by role
│   ├── layouts/       # header.ejs, footer.ejs
│   ├── client/
│   ├── freelancer/
│   ├── admin/
│   └── shared/
├── public/            # CSS, JS, images, uploads
└── database/          # schema.sql, migrate.js, seed.js
\end{lstlisting}

% ─────────────────────────────────────────────────────────────────────────────
\section{Database Design}

\subsection{Entity Overview}

The MySQL 8 schema contains 18 tables. Key entities and their relationships are described below.

\begin{table}[H]
\centering
\caption{Core database tables}
\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Table} & \textbf{Purpose} \\
\midrule
\texttt{users} & All accounts; \texttt{role ENUM('client','freelancer','admin')}, \texttt{is\_active}, \texttt{password\_hash} \\
\texttt{categories} & Job categories with icon and slug \\
\texttt{skills} & Skill tags linked to categories \\
\texttt{user\_skills} & Many-to-many: user $\leftrightarrow$ skill \\
\texttt{jobs} & Posted jobs with budget range, level, status \\
\texttt{proposals} & Freelancer bids; unique constraint on (job\_id, freelancer\_id) \\
\texttt{contracts} & Created when a proposal is accepted; tracks agreed budget \\
\texttt{deliverables} & Files/messages submitted per contract \\
\texttt{milestones} & Ordered checkpoints on a contract with approval workflow \\
\texttt{change\_orders} & Budget/deadline extension requests from freelancer \\
\texttt{messages} & Chat messages linked to a proposal OR contract \\
\texttt{notifications} & In-app notifications with read/unread state \\
\texttt{reviews} & 1--5 star reviews on completed contracts \\
\texttt{disputes} & Raised against active contracts with category and admin note \\
\texttt{offers} & Direct client-to-freelancer job offers \\
\texttt{saved\_jobs} & Freelancer bookmarks \\
\texttt{portfolio\_items} & Freelancer portfolio entries (up to 8 per user) \\
\texttt{password\_resets} & Single-use hashed tokens with 1-hour expiry \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Key Constraints and Indexes}

\begin{itemize}[noitemsep]
  \item All foreign keys use \texttt{ON DELETE CASCADE} so orphan rows are never left behind.
  \item \texttt{users.email} has a unique index; duplicate registration is rejected at the DB layer even if validation is bypassed.
  \item \texttt{proposals} has a composite unique key on \texttt{(job\_id, freelancer\_id)} to prevent duplicate bids.
  \item \texttt{jobs} is indexed on \texttt{status}, \texttt{category\_id}, and \texttt{client\_id} for fast filtered queries.
  \item \texttt{messages} is indexed on \texttt{proposal\_id} and \texttt{contract\_id} for fast chat loads.
\end{itemize}

% ─────────────────────────────────────────────────────────────────────────────
\section{Feature Implementation}

\subsection{Criterion 1--3: Core Functionality}

All forms, buttons, and interactive elements respond correctly without server errors. The three core data flows are:

\paragraph{Job $\rightarrow$ Proposal $\rightarrow$ Contract:}
A client posts a job; freelancers browse and apply with a cover letter and bid; the client reviews proposals and accepts one, which atomically creates a \texttt{contracts} row and changes the proposal status to \texttt{accepted}. All other proposals for the same job are automatically declined.

\paragraph{Contract $\rightarrow$ Milestone $\rightarrow$ Review:}
The freelancer adds milestones and marks them done; the client approves or requests revision (freelancer can re-submit). When the client marks the contract complete, both parties are prompted for a mutual review.

\paragraph{CRUD operations:}
Full create, read, update, and delete is implemented for: users (admin), jobs (client), proposals, milestones, deliverables, portfolio items, and saved jobs.

\subsection{Criterion 4--7: Password Encryption \& Security}

\subsubsection{Bcrypt Hashing (Criterion 4)}

Passwords are hashed with \texttt{bcryptjs} at cost factor 12 before any database write:

\begin{lstlisting}[language=JavaScript, caption={Password hashing on signup}]
const bcrypt = require('bcryptjs');

// Inside auth.controller.js signup handler:
const hash = await bcrypt.hash(req.body.password, 12);
await db.query(
  'INSERT INTO users (name, email, password_hash, role) VALUES (?,?,?,?)',
  [name, email, hash, role]
);
\end{lstlisting}

The column is named \texttt{password\_hash} to make its purpose explicit. The plaintext value is never stored, logged, or returned in any query result (Criterion 5).

\subsubsection{Secure Comparison (Criterion 6)}

Login uses \texttt{bcrypt.compare()} --- a constant-time comparison function --- never string equality:

\begin{lstlisting}[language=JavaScript, caption={Secure password comparison on login}]
const [rows] = await db.query(
  'SELECT * FROM users WHERE email=? AND is_active=1', [email]
);
if (!rows.length) { /* user not found */ }
const valid = await bcrypt.compare(req.body.password, rows[0].password_hash);
if (!valid) { /* wrong password */ }
\end{lstlisting}

\subsubsection{Token-Based Password Reset (Criterion 7)}

\begin{enumerate}[noitemsep]
  \item User submits their email; a 32-byte random token is generated with \texttt{crypto.randomBytes(32)}.
  \item The token is hashed with bcrypt before being stored in \texttt{password\_resets} alongside a 1-hour expiry timestamp.
  \item A link containing the \emph{raw} token is emailed to the user via Nodemailer.
  \item When the link is followed, the raw token is compared against the stored hash using \texttt{bcrypt.compare()}.
  \item On successful reset the token row is deleted (single-use guarantee).
\end{enumerate}

\subsection{Criterion 8--12: Role-Based Access Control}

\subsubsection{Three Roles (Criterion 8)}

The \texttt{users.role} column is an \texttt{ENUM('client','freelancer','admin')} enforced at the DB level. Role is written once on signup and can only be changed by an admin.

\subsubsection{Admin Route Protection (Criterion 9)}

All \texttt{/admin/*} routes pass through two middleware layers:

\begin{lstlisting}[language=JavaScript, caption={Admin route guard}]
// routes/admin.routes.js
const { requireLogin, requireRole } = require('../middleware/auth');
router.use(requireLogin, requireRole('admin'));
\end{lstlisting}

Non-admin users receive a rendered \texttt{403.ejs} page; unauthenticated users are redirected to \texttt{/login} with the intended URL saved in session for post-login redirect.

\subsubsection{Admin User Management (Criterion 10)}

The admin dashboard (\texttt{/admin/users}) lists all users with:
\begin{itemize}[noitemsep]
  \item Activate / Deactivate toggle (sets \texttt{is\_active} flag; deactivated users cannot log in).
  \item Role change dropdown (client $\leftrightarrow$ freelancer $\leftrightarrow$ admin).
  \item Account deletion with cascade cleanup.
\end{itemize}

\subsubsection{Dynamic Navigation (Criterion 11)}

\texttt{header.ejs} renders a different link set depending on \texttt{sessionUser.role}:
\begin{itemize}[noitemsep]
  \item \textbf{Client:} Dashboard · Post a Job · My Jobs · Contracts · Offers · Messages
  \item \textbf{Freelancer:} Dashboard · Browse Jobs · Contracts · Applications · Offers · Messages
  \item \textbf{Admin:} Dashboard · Users · Disputes · Reports
  \item \textbf{Guest:} Home · About · Contact · Sign in · Get started
\end{itemize}

\subsubsection{Middleware Route Guards (Criterion 12)}

Every protected route uses at least one guard:

\begin{lstlisting}[language=JavaScript, caption={Route-level access control example}]
// routes/client.routes.js
router.get('/dashboard', requireLogin, requireRole('client'), ctrl.dashboard);
router.get('/post-job',  requireLogin, requireRole('client'), ctrl.postJobForm);

// routes/freelancer.routes.js
router.get('/browse',    requireLogin, requireRole('freelancer'), jobsCtrl.browse);

// routes/contracts.routes.js — accessible to both roles
router.post('/:id/milestones',
  requireLogin, requireRole('freelancer'), ctrl.addMilestone);
router.post('/:id/milestones/:msId/review',
  requireLogin, requireRole('client'), ctrl.reviewMilestone);
\end{lstlisting}

\subsection{Criterion 13--15: Form Validation}

\subsubsection{Client-Side Validation (Criterion 13)}

\texttt{public/js/validate.js} enforces in the browser:
\begin{itemize}[noitemsep]
  \item Required field checks with red inline error messages.
  \item Email format via \texttt{type="email"} and a regex check.
  \item Password strength meter (length, uppercase, digit indicators).
  \item Budget cross-validation: max must be $\geq$ min before the form submits.
\end{itemize}

\subsubsection{Server-Side Validation (Criterion 14)}

\texttt{middleware/validate.js} exports \texttt{express-validator} rule chains for every form:

\begin{lstlisting}[language=JavaScript, caption={Signup validation rules (excerpt)}]
exports.signupRules = [
  body('name').trim().notEmpty().isLength({ max: 120 }).escape(),
  body('email').isEmail().normalizeEmail(),
  body('password')
    .isLength({ min: 8 })
    .matches(/[A-Z]/)
    .matches(/[0-9]/),
  body('role').isIn(['client', 'freelancer']),
];
\end{lstlisting}

Even if JavaScript is disabled or the request is crafted manually (e.g., via curl), the server rejects invalid inputs and never writes unsanitised data to the database.

\subsubsection{Inline Error Messages (Criterion 15)}

\texttt{validationResult(req)} extracts errors and passes them to the view. Each error is displayed inline next to the relevant field --- not as a generic alert at the top of the page.

\subsection{Criterion 16--18: Navigation \& Structure}

\subsubsection{Working Navbar (Criterion 16)}

The shared \texttt{header.ejs} layout renders a sticky navbar on every page. All links are live: no 404s, no placeholder hrefs. The notification bell polls \texttt{/api/notifications/count} every 30 seconds and shows a live badge.

\subsubsection{Logical Page Hierarchy (Criterion 17)}

The site has a clear guest-facing hierarchy (Home $\to$ About $\to$ Contact $\to$ Sign Up) and a role-specific authenticated hierarchy (Dashboard $\to$ domain-specific lists $\to$ detail pages $\to$ actions). Breadcrumb-style back links (e.g., ``$\leftarrow$ My jobs'') appear on detail pages.

\subsubsection{Sticky/Responsive Navbar \& Dropdowns (Criterion 18)}

\begin{itemize}[noitemsep]
  \item The nav bar uses \texttt{position: sticky; top: 0} with a backdrop blur.
  \item An avatar dropdown (\texttt{.nav-dropdown}) appears on click: profile link, dashboard link, sign out.
  \item A hamburger button (\texttt{\#nav-hamburger}) toggles a full-width mobile menu.
\end{itemize}

\subsection{Criterion 19--20: UI/UX Design}

\subsubsection{Design System (Criterion 19)}

All visual decisions are centralised in \texttt{public/css/style.css} using CSS custom properties:

\begin{lstlisting}[language=CSS, caption={CSS design token excerpt}]
:root {
  --brand:   #E87722;   /* Knack orange */
  --ink:     #1a1a1a;
  --ink-2:   #555;
  --ink-3:   #888;
  --paper:   #f7f6f3;
  --sunken:  #eeecea;
  --white:   #ffffff;
  --line:    #e2e0dc;
  --danger:  #d63031;
  --success: #00b894;
  --radius:  8px;
  --font-serif: 'Georgia', serif;
}
\end{lstlisting}

Spacing, font sizes, border radii, and colour are consistent throughout every page. Components (buttons, cards, badges, tables, form inputs) all inherit from this system.

\subsubsection{Responsive Design (Criterion 20)}

Layout adapts across breakpoints using CSS Grid (\texttt{.grid-2} collapses to one column on small screens), \texttt{flex-wrap}, and a mobile menu that replaces the desktop nav below 768 px. All pages were tested at 375 px, 768 px, and 1440 px.

\subsection{Criterion 21--22: Authentication \& Session Management}

\subsubsection{Login/Logout with Session (Criterion 21)}

On successful login the session is regenerated (\texttt{req.session.regenerate()}) to prevent session fixation attacks. The session stores \texttt{userId}, \texttt{userName}, \texttt{role}, and \texttt{avatarUrl}. Logout calls \texttt{req.session.destroy()} and clears the cookie.

\begin{lstlisting}[language=JavaScript, caption={Session creation on login}]
req.session.regenerate((err) => {
  req.session.userId    = user.id;
  req.session.userName  = user.name;
  req.session.role      = user.role;
  req.session.avatarUrl = user.avatar_url;
  res.redirect(returnTo || '/dashboard');
});
\end{lstlisting}

\subsubsection{Session Expiry (Criterion 22)}

The session cookie has \texttt{maxAge: 1000 * 60 * 30} (30 minutes). Any request after this window is unauthenticated and the user is redirected to \texttt{/login}. \texttt{httpOnly: true} prevents client JavaScript from reading the cookie.

\subsection{Criterion 23--25: Git Version Control}

\subsubsection{Repository (Criterion 23)}

\href{https://github.com/mmujtaba0085/knack-freelancer-marketplace}{github.com/mmujtaba0085/knack-freelancer-marketplace} --- public, properly structured, \texttt{.gitignore} excludes \texttt{node\_modules/}, \texttt{.env}, and \texttt{public/uploads/*}.

\subsubsection{Meaningful Commits (Criterion 24)}

The repository has \textbf{21 commits} each covering a discrete, testable feature. Commits are ordered to reflect real incremental development from scaffold through to documentation.

\subsubsection{Commit Message Convention (Criterion 25)}

All commit messages follow \href{https://www.conventionalcommits.org/}{Conventional Commits}:
\begin{itemize}[noitemsep]
  \item \texttt{chore:} --- project scaffold
  \item \texttt{feat:} --- all feature additions (database, auth, jobs, proposals, contracts, messaging, offers, notifications, disputes, etc.)
  \item \texttt{fix:} --- bug corrections
  \item \texttt{docs:} --- README and team attribution
\end{itemize}

\subsection{Criterion 26: Footer}

A consistent footer appears on every page via \texttt{views/layouts/footer.ejs}. It contains:
\begin{itemize}[noitemsep]
  \item \textbf{Brand column:} Knack logo, tagline, social links (𝕏, LinkedIn, Instagram).
  \item \textbf{Marketplace column:} Browse Jobs, Post a Job, Join as Freelancer, Hire Talent.
  \item \textbf{Company column:} About, Contact, Blog, Careers.
  \item \textbf{Legal column:} Terms of Service, Privacy Policy, Cookie Policy.
  \item \textbf{Footer bar:} Dynamic copyright year (\texttt{new Date().getFullYear()}) and build credit.
\end{itemize}

\subsection{Criterion 27--32: Content \& Creativity}

\subsubsection{Original Content (Criterion 27)}

All page copy, job descriptions, freelancer bios, and UI microcopy were written specifically for Knack. The seed data contains 33 realistic users (with names, headlines, bios, and locations), 40 varied job listings across 8 categories, and 97 proposals with bespoke cover letters.

\subsubsection{Icons \& Multimedia (Criterion 28)}

\begin{itemize}[noitemsep]
  \item SVG default avatar used throughout; replaced by user-uploaded JPEG/PNG avatars.
  \item Unicode emoji used tastefully as icons (✦, 💬, ★, 📅, 💰) --- no external icon library dependency.
  \item Chat view renders uploaded images inline.
\end{itemize}

\subsubsection{Innovative Concept (Criterion 29)}

Knack goes well beyond a CRUD tutorial. Unique features include:
\begin{itemize}[noitemsep]
  \item \textbf{Proposal-stage chat} --- client and freelancer can discuss requirements before the contract starts.
  \item \textbf{Milestone approval workflow} --- granular deliverable sign-off with revision notes.
  \item \textbf{Change orders} --- structured budget/timeline amendment requests within a live contract.
  \item \textbf{Direct offers} --- client initiates a contract from a freelancer's profile without a public job posting.
  \item \textbf{Rate-limited message notifications} --- one notification per 5-minute window per thread to prevent spam.
  \item \textbf{Re-hire shortcut} --- one-click re-engagement flow on completed contracts.
\end{itemize}

\subsubsection{Visual Design Quality (Criterion 30)}

\begin{itemize}[noitemsep]
  \item \textbf{Colour:} A single brand orange (\texttt{\#E87722}) paired with near-black ink and warm paper tones.
  \item \textbf{Typography:} System sans-serif for body copy; Georgia serif for numeric values and headings to create visual contrast.
  \item \textbf{Whitespace:} 24 px base spacing unit used consistently; cards use 24 px internal padding.
  \item \textbf{Badges:} Colour-coded status chips (\texttt{.badge-open}, \texttt{.badge-accepted}, \texttt{.badge-rejected}, etc.).
\end{itemize}

\subsubsection{Animations \& Micro-interactions (Criterion 31)}

\begin{itemize}[noitemsep]
  \item \textbf{Toast notifications:} slide-up / fade-out animation (\texttt{@keyframes slideUp}).
  \item \textbf{Button hover:} subtle scale transform and shadow transition on all \texttt{.btn} elements.
  \item \textbf{Nav dropdown:} CSS opacity + translateY transition on \texttt{.nav-dropdown}.
  \item \textbf{Skill checkboxes:} border colour transition when a skill chip is selected.
  \item \textbf{Star picker:} colour transition on hover using \texttt{:has()} selector.
  \item \textbf{Chat bubble:} \texttt{scroll-behavior: smooth} on container; new messages scroll into view.
\end{itemize}

\subsubsection{Creative Impression (Criterion 32)}

The Knack aesthetic is intentionally premium: restrained colour palette, editorial serif numerics, generous whitespace, and copy that avoids generic placeholder text. The ✦ motif (used as favicon, empty-state icon, and decorative element) gives the brand a distinctive identity. The overall feel resembles a real SaaS product rather than a student project.

\subsection{Criterion 33: Performance \& Optimisation}

\begin{itemize}[noitemsep]
  \item The single \texttt{style.css} file is hand-authored with no unused frameworks or libraries bundled in.
  \item No external JavaScript frameworks are loaded --- all interactivity is Vanilla JS.
  \item MySQL queries use parameterised prepared statements (preventing SQL injection and enabling query plan caching).
  \item Database indexes on \texttt{jobs.status}, \texttt{jobs.category\_id}, \texttt{messages.contract\_id}, etc., ensure sub-millisecond lookups on seeded data.
  \item Avatar images are processed through \texttt{multer} to enforce a maximum 10 MB upload cap.
  \item The notification polling endpoint (\texttt{/api/notifications/count}) returns a single integer JSON response to minimise bandwidth.
\end{itemize}

\subsection{Criterion 34--35: Documentation \& Presentation}

\subsubsection{README (Criterion 34)}

A comprehensive \texttt{README.md} is included in the repository root. It covers:
\begin{itemize}[noitemsep]
  \item Team members and roll numbers.
  \item Prerequisites (Node.js 18+, MySQL 8+).
  \item Step-by-step setup: clone, \texttt{npm install}, \texttt{.env} configuration, schema import, seed, start server.
  \item Test account credentials for all three roles.
  \item Annotated project directory tree.
  \item Security implementation summary.
\end{itemize}

\subsubsection{Walkthrough (Criterion 35)}

A live demo will be provided during the presentation covering the full user journey: guest home $\to$ signup $\to$ post job $\to$ apply $\to$ accept proposal $\to$ add milestones $\to$ submit deliverable $\to$ complete contract $\to$ review $\to$ admin panel.

% ─────────────────────────────────────────────────────────────────────────────
\section{Self-Evaluation}

\begin{longtable}{p{0.5cm} p{9cm} p{1.8cm} p{1.8cm}}
\caption{Self-Evaluation Sheet --- Knack Project}\\
\toprule
\textbf{Sr\#} & \textbf{Criterion} & \textbf{Obtained} & \textbf{Total} \\
\midrule
\endfirsthead
\multicolumn{4}{c}{\small\textit{(continued)}}\\
\toprule
\textbf{Sr\#} & \textbf{Criterion} & \textbf{Obtained} & \textbf{Total} \\
\midrule
\endhead
\bottomrule
\endfoot

\multicolumn{4}{l}{\textbf{Functionality}} \\[4pt]
1 & All core features implemented correctly (forms, buttons, interactive elements respond without errors) & 10 & 10 \\[4pt]
2 & Login and Signup system working end-to-end (registration, authentication, session start) & 10 & 10 \\[4pt]
3 & Data processing functional (database CRUD or API integration working correctly) & 5 & 5 \\[8pt]

\multicolumn{4}{l}{\textbf{Password Encryption \& Security}} \\[4pt]
4 & Passwords hashed using bcrypt before storing in database & 8 & 8 \\[4pt]
5 & Plain-text passwords never stored or logged anywhere in the system & 5 & 5 \\[4pt]
6 & Password comparison done securely using hash-comparison functions (not string equality) & 4 & 4 \\[4pt]
7 & Password reset flow implemented securely (token-based, time-limited link) & 3 & 3 \\[8pt]

\multicolumn{4}{l}{\textbf{Role-Based Access Control (Admin \& User)}} \\[4pt]
8  & At least two distinct roles defined: Admin and Regular User, stored in database & 5 & 5 \\[4pt]
9  & Admin dashboard accessible only to admin role; regular users redirected or shown 403 & 8 & 8 \\[4pt]
10 & Admin can manage users: view list, activate/deactivate accounts, change roles & 7 & 7 \\[4pt]
11 & Frontend navigation dynamically changes based on logged-in user role & 5 & 5 \\[4pt]
12 & Backend routes protected with middleware/guard that checks role before processing & 5 & 5 \\[8pt]

\multicolumn{4}{l}{\textbf{Form Validation}} \\[4pt]
13 & Client-side validation: required fields, email format, password strength enforced in browser & 5 & 5 \\[4pt]
14 & Server-side validation: inputs sanitized and validated before processing & 5 & 5 \\[4pt]
15 & Clear, inline error messages displayed to user for invalid or missing inputs & 5 & 5 \\[8pt]

\multicolumn{4}{l}{\textbf{Navigation \& Structure}} \\[4pt]
16 & Working navbar on every page; all links accessible with no broken routes & 5 & 5 \\[4pt]
17 & Logical page hierarchy (Home, About, Contact, etc.) with smooth user flow & 3 & 3 \\[4pt]
18 & Sticky/responsive navbar, dropdown menus, breadcrumbs, or site search & 2 & 2 \\[8pt]

\multicolumn{4}{l}{\textbf{UI / UX Design}} \\[4pt]
19 & Clean, consistent layout --- uniform colors, fonts, and spacing throughout & 5 & 5 \\[4pt]
20 & Responsive design: layout adapts correctly on desktop, tablet, and mobile screens & 5 & 5 \\[8pt]

\multicolumn{4}{l}{\textbf{Authentication \& Session Management}} \\[4pt]
21 & Functional login/logout with correct session handling and cookie use & 10 & 10 \\[4pt]
22 & Session expires after inactivity; re-login required for sensitive actions & 5 & 5 \\[8pt]

\multicolumn{4}{l}{\textbf{Git Version Control}} \\[4pt]
23 & Repository initialized and hosted on GitHub with proper project structure & 3 & 3 \\[4pt]
24 & Minimum 10 meaningful commits reflecting incremental development progress & 4 & 4 \\[4pt]
25 & Clear commit messages following a consistent naming convention & 3 & 3 \\[8pt]

\multicolumn{4}{l}{\textbf{Footer \& Layout Components}} \\[4pt]
26 & Footer present on all pages with contact info, social links, and copyright notice & 5 & 5 \\[8pt]

\multicolumn{4}{l}{\textbf{Content \& Creativity}} \\[4pt]
27 & Original, relevant, and well-written content across all pages & 5 & 5 \\[4pt]
28 & Effective use of images, icons, illustrations, or multimedia that enhance the design & 5 & 5 \\[4pt]
29 & Unique or innovative project concept --- goes beyond a generic template or tutorial & 5 & 5 \\[4pt]
30 & Visual design quality: color palette, typography, whitespace, and overall aesthetic polish & 5 & 5 \\[4pt]
31 & Animations, transitions, or micro-interactions used tastefully to enhance UX & 5 & 5 \\[4pt]
32 & Overall creative impression: layout creativity, theme consistency, and user delight factor & 5 & 5 \\[8pt]

\multicolumn{4}{l}{\textbf{Performance \& Optimization}} \\[4pt]
33 & Pages load quickly; images and assets optimized; minimal unnecessary scripts or stylesheets & 5 & 5 \\[8pt]

\multicolumn{4}{l}{\textbf{Documentation \& Presentation}} \\[4pt]
34 & README or project report included with clear explanation of features and setup steps & 3 & 3 \\[4pt]
35 & Live demo or recorded walkthrough provided during presentation & 2 & 2 \\[8pt]

\midrule
\multicolumn{2}{r}{\textbf{Total}} & \textbf{180} & \textbf{180} \\
\bottomrule
\end{longtable}

% ─────────────────────────────────────────────────────────────────────────────
\section{Conclusion}

Knack demonstrates a production-quality full-stack Node.js application implementing all 35 rubric criteria. The codebase is organised, secure, and maintainable: strict role-based access control, bcrypt security throughout the auth flow, parameterised SQL queries, server-side validation on every form, and a clean MVC architecture with one controller and one route file per domain area. Beyond the rubric requirements, the project delivers advanced features --- milestone-driven contract management, real-time-like messaging with file sharing, direct offers, change orders, and rate-limited notifications --- that make Knack a genuinely useful freelancer marketplace rather than a textbook exercise.

\vspace{1cm}
\begin{center}
  \rule{0.6\linewidth}{0.4pt}\\[0.3cm]
  {\large\textbf{\textcolor{brand}{Knack}}}\\
  {\small The marketplace for serious work.}\\[0.2cm]
  {\small Muhammad Mujtaba (22i-2123) \quad Shahmeer (23i-2063)}\\[0.2cm]
  {\small \href{https://github.com/mmujtaba0085/knack-freelancer-marketplace}{github.com/mmujtaba0085/knack-freelancer-marketplace}}
\end{center}

\end{document}
%   m i n o r   f o r m a t t i n g   r e v i e w  
 %   f i n a l   f o r m a t t i n g   c h e c k  
 